---
title: "Montecarlo Estimation"
author: "Luis Avila"
date: "7/28/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####A simple MonteCarlo approximation in R

####We are going to estimate $\pi$.

Imagine a circle in a square of side=1. The radio of that circle would be r=0.5.

We know the area of a circle is $$\pi*r^2$$ 
and the area of the square in terms of the radio would be $$(2*r)^2$$

Now, a ratio of those areas would be
$$\rho=\frac{Area of Circle}{Area of Square}=\frac{\pi*r^2}{(2*r)^2}=\frac{\pi}{4}$$

so that $$\pi=4*\rho$$

Now, to approximate $\rho$ we will generate random points in [0-1,0-1] and calculate the ratio of points that fall into the circunference to the points that fall off the circuference.

Calculating for 100 points

```{r}
 n<-100
 x<-runif(n,0,1)
 y<-runif(n,0,1)
 d<-(x^2+y^2<1)
 plot(x,y,col=d+1,pch=19)
 
 in.the.circle<-sum(d)
 in.the.circle
 off.the.circle<-(n-in.the.circle)
 off.the.circle

 pi.100<-(in.the.circle/n)*4 
 pi.100

```

We repeat for 1,000 points
```{r}
 n<-1000
 x<-runif(n,0,1)
 y<-runif(n,0,1)
 d<-(x^2+y^2<1)
 plot(x,y,col=d+1,pch=19)
 
 in.the.circle<-sum(d)
 in.the.circle
 off.the.circle<-(n-in.the.circle)
 off.the.circle

 pi.1000<-(in.the.circle/n)*4 
 pi.1000

```

We repeat for 10,000 points
```{r}
 n<-10000
 x<-runif(n,0,1)
 y<-runif(n,0,1)
 d<-(x^2+y^2<1)
 plot(x,y,col=d+1,pch=19)
 
 in.the.circle<-sum(d)
 in.the.circle
 off.the.circle<-(n-in.the.circle)
 off.the.circle

 pi.10000<-(in.the.circle/n)*4 
 pi.10000

```
And for 200,000 points
```{r}
 n<-200000
 x<-runif(n,0,1)
 y<-runif(n,0,1)
 d<-(x^2+y^2<1)
 #plot(x,y,col=d+1,pch=19)
 
 in.the.circle<-sum(d)
 in.the.circle
 off.the.circle<-(n-in.the.circle)
 off.the.circle

 pi.200000<-(in.the.circle/n)*4 
 pi.200000

```
We can see that our approximation of $\pi$ tends to get closer to the actual `r pi` as we increase the number of randomly generated points.